Very simplified likelihood

### 2024-01-21 inst/very_simplified_likelihood.Rmd
further work

Probability of the focal competitor coming first:

\[
\frac{a}{a+nb}=\frac{a}{n - a(n-1)a}=\frac{1-b}{1+(n-1)b}
\]

Coming second:
\[
\frac{b}{a+nb}\cdot\frac{a}{a+(n-1)b}
\]

Third:

\[
\frac{b}{a+nb}\cdot\frac{b}{a+(n-1)b}\cdot\frac{a}{a+(n-2)b}
\]

### Likelihood functions

@@

\[\mathcal{L}(a)=\frac{a}{a+nb}=\frac{a}{n+(n-1)a}\]

@@

\[
n=2\longrightarrow\mathcal{L}=\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^2}{1+b}
\]

\[
n=3\longrightarrow\mathcal{L}=\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^3}{(1+2b)(1+b)}
\]

\[
n=4\longrightarrow\mathcal{L}=\frac{b}{a+4b}\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=
\frac{b^4}{(1+3b)(1+2b)(1+b)}
\]

@@

\[
\frac{1}{(B+n)(B+n-1)\cdots(B+1)B}=\frac{(B-1)!}{(B+n)!}=\frac{\Gamma(B)}{\Gamma(B+n+1)}
\]

@@

Now try $r=3$, 4 competitors.   Taking things one step at a time:

\[
\mbox{prob of coming first}      =
3!\cdot\frac{a}{a+3b}\cdot\frac{b}{3b}\cdot\frac{b}{2b}\cdot\frac{b}{b}
\]

@@

>>>>

@@

\[
\mathcal{L}(a)=\frac{a}{a+nb}=\frac{1-b}{1+(n-1)b}
\]

@@

\[
n=2\longrightarrow\mathcal{L}=\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^2}{1+b}
=\frac{1}{B(B+1)}
\]

\[
n=3\longrightarrow\mathcal{L}=\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=\frac{b^3}{(1+2b)(1+b)}=\frac{1}{B(B+1)(B+2)}
\]

\[
n=4\longrightarrow\mathcal{L}=\frac{b}{a+4b}\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}=
\frac{b^4}{(1+3b)(1+2b)(1+b)}=
\frac{1}{B(B+1)(B+2)(B+3)}
\]

@@

\[
\frac{1}{(B+n-1)(B+n-2)\cdots(B+1)B}=\frac{(B-1)!}{(B+n)!}=\frac{\Gamma(B)}{\Gamma(B+n+1)}
\]

@@

Now try $r=3$ [i.e. 4 competitors].  Taking things one step at a time:

\[
\mbox{prob of coming first} =
3!\cdot\frac{a}{a+3b}\cdot\frac{b}{3b}\cdot\frac{b}{2b}\cdot\frac{b}{b}=
\frac{1-b}{1+2b}
\]

\[
\mbox{prob of coming second} =
3!\cdot\frac{b}{a+3b}\cdot\frac{a}{a+2b}\cdot\frac{b}{2b}\cdot\frac{b}{b}=
3\cdot\frac{b(1-b)}{(1+2b)(1+b)}
\]

\[
\mbox{prob of coming third} =
3!\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{a}{a+b}\cdot\frac{b}{b}
=
6\cdot\frac{b^2(1-b)}{(1+2b)(1+b)}
\]
\[
\mbox{prob of coming fourth (last)} =
3!\cdot\frac{b}{a+3b}\cdot\frac{b}{a+2b}\cdot\frac{b}{a+b}\cdot\frac{a}{a}
=
6\cdot\frac{b^3}{(1+2b)(1+b)}
\]

The sum would be $[(1+b)(1-b) + 3b(1-b) + 6b^2(1-b) +
6b^3]/[(1+b)(1+2b)]=1$.

Now consider the general case.  Again, $n$ clones of strength $b$ and
one of strength $a=1-b$, $n+1$ competitors.

\[
\mbox{prob of coming first} =
n!\cdot\frac{b}{a+nb}\cdot\frac{b}{nb}\cdot\frac{b}{(n-1)b}\cdots\frac{b}{b}
=
\frac{b}{1+(n-1)b}
\]

\[
\mbox{prob of coming second} =
n!\cdot\frac{b}{a+nb}\cdot\frac{a}{a+(n-1)b}\cdot\frac{b}{(n-1)b}\cdot\frac{b}{(n-2)b}\cdots\frac{b}{2b}\cdot\frac{b}{b}
=n\cdot X
\]

@@

## Masterchef

Look at the dotchart in the _MasterChef_ part of the `hyper2`
vignette.  We could use a Very Simplified Likelihood to estimate, say,
Laura's BT strength assuming that the other competitors are clones.
